### 2018-09-17 notes/predict.tex

\documentclass[a4paper, 11pt]{article}
\usepackage{fullpage, amssymb, amsmath, graphicx, bm}

\newcommand{\mytitle}{}

\begin{document}
\noindent
\large\textbf{\mytitle} \\ \\ Tyler Gordon \\
\normalsize \today 
\ \ \hrulefill
\section{Input coordinates in second dimension are same as data coordinates}
In the 1D case, the predictive distribution of a GP at coordinates $\bm{y^*} = (t^*_1, t^*_2... t^*_p)^T$ has a mean given by 
\begin{equation}
	\mu_p^* = \mu_\theta(t_p^*) + K(t_p^*, \bm{t})\bm{z}
\end{equation}
where $\bm{z} = K(\bm{t}, \bm{t})^{-1}[\bm{y} - \bm{\mu_\theta}(\bm{t})]$. In the 2D case for a GP with covariance in the second dimension 
defined the covariance matrix $Q(\bm{u},\bm{u})$, the predictive distribution is given by 
\begin{equation}
	\bm{\mu_p}^* = \bm{\mu_\theta}(t_p^*) + K(t_p^*, \bm{t})X
\end{equation}
where $X_p = Q(\bm{u}, \bm{u})\bm{z}_{pM:(p+1)M}$ is a vector of length M and 
$\bm{z} = K_{2D}^{-1}[\bm{y} - \bm{\mu_\theta}(\bm{t})]$ where $K_{2D}$ is the full 
2D covariance matrix and $\bm{y} = [y(t_1, u_1), y(t_1, u_2), ...y(t_1, u_M), y(t_2, u_1), ...y(t_N, u_M)]^T$ is the dataset on which 
the GP is conditioned. The algorithm for computing the predictive distribution is given by equations B26-B30 in FM17 with the 
substitution of $X$ for $\bm{z}$. The result is a matrix $\mu^* \in \mathbb{Z}^{N x M}$ where $\mu_{n, m}$ is the predicted mean 
at coordinates $(t_n, u_q)$. 
\section{Input coordinates in second dimension are different from data coordinates}
When the input coordinates for the second dimension do not coincide with the data coordinates, the matrix X becomes 
\begin{equation}
	X_p = Q(\bm{u}^*,\bm{u})\bm{z}_{pM:(p+1)M}
\end{equation}
From which we see that $\bm{X}_p$ is just the mean of the predictive distribution for a 1D process with kernel Q at input coordinates $\bm{u}^*$. 
If the covariance in the second dimension is also defined by a celerite kernel, the cost of computing each $\bm{X}_p$ is  
$\mathcal{O}(mM + rR)$ where $M$ is the number of data coordinates in the second dimension, $R$ is the number of 
input coordinates, and $m$ and $r$ are some constants. Therefore the total cost of computing $X$ is $\mathcal{O}(mMP + rRP)$, 
and the total cost of the 2D prediction algorithm is $\mathcal{O}(mMP + rRP + pP + nN) \sim \mathcal{O}((mM+rR+p)P + nN)$
\end{document}